Secure Composition of Robust and Optimising Compilers

### 2023-07-09 popl/main.tex

The specification trace model defines events as either the empty event ($\emptyevent$), a crash ($\lightning$), or as tuples consisting of a pre-event, a control-tag, and a security-tag.
The purpose of the model is to define properties, such as a stricter variant of cryptographic constant time.
To this end, security-tags indicate whether an event contains sensitive information ($\lock$) or not ($\unlock$), while control-tags state whether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event.
The latter is necessary to be able to ignore actions done by a spurious context that, e.g., immediately deallocates a memory location twice, thus violating \gls{tms}.
%Even though this could be used for blame~\cite{patrignani2023blame}, this is not what this paper is concerned about.
Lastly, pre-events describe the actual kind of event that happened.
One such kind is the allocation event ($\ev{Alloc\ \loc\ n}$) that fires whenever a program claims $n$ cells of memory and stores them at address $\loc$.
Likewise, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
These two events alone allow already a partial description of \gls{tms} by requiring that, e.g., there is only one deallocation event that carries a location $\loc$.
To be able to express \gls{sms}, there is also an event to describe reads from and writes to memory ($\ev{Use\ \loc\ n}$).
Finally, for cryptographic code, there is a general guideline that secrets must not be visible on a trace.
Moreover, an instruction whose timing is data-dependent must not have a secret as an operand.
Typical operations with data-dependent timing are branches and certain binary operations, such as division\footnote{This is highly architecture dependent, but division is an operation that serves as a classic example for a data-dependent timing instruction.}.
Both operations are also modelled in the specification trace model ($\ev{Branch\ n}$ and $\ev{Binop\ n}$).

@@

%\section{Monitors and Trace Agreements\pages{1.5}}\label{sec:monitors}
\subsection{Monitors\pages{1.5}}\label{sec:monitors}
This section introduces monitors for each of the properties presented earlier (\Cref{subsec:propdefs}).
The use of monitors is because they have some structure useful for inductive proofs, whereas traces lack this structure, thus making proofs difficult.
To lessen the burden when proving, each monitor makes use of a custom trace model that contains only the relevant information related to the property the monitor checks.
To project specification traces $\trace$ onto these monitor-level traces $\abstrace$, each property has an associated event agreement relation $\traceagree[]{\pi}{\trace}{\abstrace}$.
This relation performs the actual work of projecting the events, while the stutter-equivalence $\traceagree{\pi}{\trace}{\abstrace}$ remains the same for all properties $\pi$, yet is indexed by $\pi$ to emphasize the kind of event agreement.
With agreements, this section starts to define monitor satisfaction for traces and states that monitor satisfaction implies property satisfaction.
To this end, monitor satisfaction is defined as follows.
\bul{A specification trace $\trace$ monitor-satisfies property $\pi$} iff there exists a (final) monitor state $\monitor$ and an abstract trace $\abstrace$ such that the \iul{specification trace $\trace$ agrees with abstract trace $\abstrace$} and \oul{the initial monitor\footnote{In this paper, for all monitors, the initial monitor state is denoted as $\emptyset$.} can step to the (final) monitor state $\monitor$ with abstract trace $\abstrace$}.

@@

Notably, real-world compilation chains also perform a series of passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
Both examples can be seen in the wild~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion}, it is infeasible to list them all.
Consider the compilation passes of LLVM\cite{lattner2004llvm} whose source-code optimisations are industry-competitive.
In LLVM, the pass manager can be used to schedule passes in a way where the optimisations yield the best results.
Take \gls{cf} and \gls{dce}.
Unfortunately, the order in which \gls{cf} and \gls{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
This {\em phase ordering problem} is well-known in literature and a practical solution is to simply perform a fixpoint iteration.
Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
The following corollary justifies that any such order of compilation passes is valid with respect to security.
So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.

@@

Besides the sequential composition, there are two other compositions that are more at the language- than the compiler-level.
Nevertheless, consider $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
As an example, $\src{Java}$ and $\obj{Kotlin}$ compile to $\trg{JVM BC}$.
Technically speaking, they both lower to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language, or rather, intermediate representation, performs a kind of {\em upper} composition.
Given a secure compiler from $\src{Java}$ to $\trg{JVM BC}$ and one from $\obj{Kotlin}$ to $\trg{JVM BC}$, it is easy to envision that one can engineer a combined compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$.
A practical example of this, which is, however, insecure, is {\tt Android Studio}~\cite{androidstudio}.
The proof of \Cref{thm:rtp} can be easily adapted to also works for this {\em upper} composition.
The corresponding theorem is not listed, since there is no general definition of this kind of composition.

@@

\begin{figure}
  \begin{center}
    \begin{tikzpicture}
      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left] (progunopt) {\begin{lstlisting}[language=ml,basicstyle=\scriptsize\ttfamily]
let a = true in
if a then
  print "a"
else
  print "b"
\end{lstlisting}};
      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progunopt] (progoptcf) {\begin{lstlisting}[language=ml,basicstyle=\scriptsize\ttfamily]
if true then
  print "a"
else
  print "b"
\end{lstlisting}};
      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progoptcf] (progoptcfdce) {\begin{lstlisting}[language=ml,basicstyle=\scriptsize\ttfamily]
print "a"
\end{lstlisting}};% the following is a hack to get my syntax highlighting to work again \end{lstlisting}
      % edges
      \draw[->,very thick] (progunopt) edge[loop left,left] node {\gls{dce}} (progunopt);
      \draw[->,very thick] (progunopt) edge[sloped,above] node {\gls{cf}} (progoptcf);
      \draw[->,very thick] (progoptcf) edge[sloped,above] node {\gls{dce}} (progoptcfdce);
    \end{tikzpicture}
  \end{center}
  \caption{Example program where the level of optimisations differ for one pass of applying \gls{cf} and \gls{dce} in any order. %
  Every edge is a compilation pass and the label on the edge states what the pass does, i.e., \gls{cf} or \gls{dce}. %
  The source code in the nodes is a glorified compiler intermediate representation and the code gets more optimised towards the righthandside of the figure.}\label{fig:cfdceex}
\end{figure}

%\section{Case Study: Memory Safe and Cryptographic Constant Time Code}\label{sec:casestud}
\section{Case Study: Preliminary Definitions\pages{2}}\label{sec:casestud:defs}

This section defines programming languages that the secure compilers defined in the next section will make use of.
To this end, this section defines the languages $\Ltms$, $\Ltrg$, $\Lms$, and $\Lscct$.
In this paper, $\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls{tms}.
However, $\Ltms$ programs are not \gls{sms}.
That is, there are well-typed $\Ltms$ programs that perform an out-of-bounds access.
In contrast, $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls{ms}.
$\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability.
All three languages -- so $\Ltms$, $\Ltrg$, and $\Lms$ -- assume \gls{cct} to hold.
Even though it is possible to write programs that, e.g., return early in loops, which is trivially not \gls{cct}, they can be transformed to adhere to \gls{cct}~\cite{cauligi2019fact}.
The language $\Lscct$ allows violating \gls{cct} by emitting events on, e.g., branching and division, that contain secrets.
As a countermeasure, the language provides a way to read and write to a {\em model-specific register} that enables a ``\gls{cct}-mode''.

\subsection{Shared Language Definitions}\label{subsec:cs:defs}

\begin{gather*}
  \begin{aligned}
  \mi{(Expressions)}~\expr&\bnfdef\ x \mid \valueexpr \mid \lbinop{\expr[_{1}]}{\expr[_{2}]} \mid \lget{x}{\expr} \mid \lnew{\expr[_{1}]}{\expr[_{2}]} \mid \ldelete{x} \mid \lset{x}{\expr[_{1}]}{\expr[_{2}]} \\
    & \mid \llet{x}{\expr[_{1}]}{\expr[_{2}]}  \mid \lreturn{\expr} \mid \lcall{g}{\expr} \mid \lifz{\expr}{\expr[_{1}]}{\expr[_{2}]} \mid \labort \\
  \mi{(Symbols)}~&\function\bnfdef\ \lfunction{g}{x}{e}\hspace{0.5cm}
  \mi{(Libraries)}~\library\bnfdef\ \hole{\cdot} \mid \function,\library \\
  \mi{(Programs)}~&\prog{\library_{\ctx}}{\library_{\comp}}\hspace{0.5cm}
  \mi{(Substitutions)}~\substvar\bnfdef \hole{\cdot} \mid \subst{\valueexpr}{x},\gamma \\
  \end{aligned}
\end{gather*}

Above is a fraction of the shared syntax of all the programming languages of this paper.
The syntax and semantics are mostly standard.
However, when allocating pointers with $\lnew{\expr[_{1}]}{\expr[_{2}]}$, the expression $\expr[_{1}]$ determines the initial value of the objects and $\expr[_{2}]$ the overall size of an array.
The data associated to pointers resides in the dynamic state $\Omega$ that contains sandboxed heaps, one for the context and the other for the component.
Because of the sandboxing, locations pointing into heaps must be associated with a control-tag to know the heap they point into.
Other metadata in $\Omega$ entails a list of callable objects with information on whether they belong to the context or the component.
That way, when calling a callable object, the semantics can take care of emitting appropriate events that signal a context switch.
For the languages considered in this paper, no callable objects are first-class.
So, the only values are pairs, pointers, and natural numbers.
Another non-standard part of the languages is the concept of the poison-tag.
The poison-tag marks a location as freed ($\poisoned$) without removing the information associated to the freed pointer from $\Omega$.
This measure prevents the semantics from getting stuck when an already freed pointer is freed again.
Because of this, it is possible that programs like $\llet{x}{\lnew{0}{5}}{\ldelete{x};\ldelete{x}}$ reduce and emit a trace along the lines of $\ev{Alloc\ \loc\ \sandboxtag\ 5}\cdot\ev{Dealloc\ \loc\ \sandboxtag}\cdot\ev{Dealloc\ \loc\ \sandboxtag}$.

\begin{gather*}
  \begin{aligned}
  \mi{(Pre\text{-}Events)}~\event[_{b}]\bnfdef \ev{Alloc}\ \loc\ \valueexpr \mid \ev{Dealloc}\ \loc \mid \ev{Get}\ \loc\ \valueexpr &\mid \ev{Set}\ \loc\ \valueexpr\ \valueexpr['] \mid \ev{Call}\ \comm\ g\ \valueexpr \mid \ev{Ret}\ \comm\ \valueexpr \mid \ev{Start} \mid \ev{End}\ \valueexpr \\
  \mi{(Events)}~\event&\bnfdef \emptyevent \mid \lightning \mid \tup{\event[_{b}]}{\sandboxtag} \\
  \end{aligned}
\end{gather*}

>>>>

The specification trace model defines events as either the empty event ($\emptyevent$), a crash ($\lightning$), or as tuples consisting of a pre-event, a control-tag, and a security-tag.
The purpose of the model is to define key security properties of interest, such as \gls{ms} or a stricter variant of cryptographic constant time.
To this end, security-tags indicate whether an event contains sensitive information ($\lock$) or not ($\unlock$), while control-tags state whether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event.
The latter is necessary to be able to ignore actions done by a spurious context that, e.g., immediately deallocates a memory location twice, thus violating \gls{tms}~\cite{nagarakatte2010cets}.
%Even though this could be used for blame~\cite{patrignani2023blame}, this is not what this paper is concerned about.
Lastly, pre-events describe the actual kind of event that happened.
One such kind is the allocation event ($\ev{Alloc\ \loc\ n}$) that fires whenever a program claims $n$ cells of memory and stores them at address $\loc$.
Dually, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
These two events alone allow\MP{allow who? allow needs an object, this is improper english} already a partial description of \gls{tms} by requiring that, e.g., there is only one deallocation event that carries a location $\loc$.
To be able to express \gls{sms}, there is also an event to describe reads from and writes to memory ($\ev{Use\ \loc\ n}$).
Finally, for cryptographic code, there is a general guideline that secrets must not be visible on a trace.
Moreover, an instruction whose timing is data-dependent must not have a secret as an operand.
Typical operations with data-dependent timing are branches and certain binary operations, such as division\footnote{This is highly architecture-dependent, but division is an operation that serves as a classic example for a data-dependent timing instruction.}.
Both operations are also modelled in the specification trace model ($\ev{Branch\ n}$ and $\ev{Binop\ n}$).

@@

%\section{Monitors and Trace Agreements\pages{1.5}}\label{sec:monitors}
\subsection{Monitors\pages{1.5}}\label{sec:monitors}

The use of monitors is because they have some structure useful for inductive proofs, whereas traces lack this structure, thus making proofs difficult.\MP{not true. safety props are enforced by monitors}
To lessen the burden when proving, each monitor makes use of a custom trace model that contains only the relevant information related to the property the monitor checks.
To project specification traces $\trace$ onto these monitor-level traces $\abstrace$, each property has an associated event agreement relation $\traceagree[]{\pi}{\trace}{\abstrace}$.
This relation performs the actual work of projecting the events, while the stutter-equivalence $\traceagree{\pi}{\trace}{\abstrace}$ remains the same for all properties $\pi$, yet is indexed by $\pi$ to emphasize the kind of event agreement.
With agreements, this section starts to define monitor satisfaction for traces and states that monitor satisfaction implies property satisfaction.
To this end, monitor satisfaction is defined as follows.
\bul{A specification trace $\trace$ monitor-satisfies property $\pi$} iff there exists a (final) monitor state $\monitor$ and an abstract trace $\abstrace$ such that the \iul{specification trace $\trace$ agrees with abstract trace $\abstrace$} and \oul{the initial monitor\footnote{In this paper, for all monitors, the initial monitor state is denoted as $\emptyset$.} can step to the (final) monitor state $\monitor$ with abstract trace $\abstrace$}.

@@

\subsubsection{Securing Optimisations}
\MPin{note that i'm always using british spelling. optimisation VS optimization. etc. }

Notably, real-world compilation chains also perform a series of (sequential) passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
\MPin{
	do not say the negation of something before the non-negation.
	say what things are first, and what they are not later (if at all)
}
Both examples can be seen in the wild~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion}, it is infeasible to list them all\MP{survey to cite? this is a bit informal}.
Consider the compilation passes of LLVM~\cite{lattner2004llvm} whose source-code optimisations are industry--competitive\MP{meaning?}.
In LLVM, the pass manager\MP{meaning} can be used to schedule passes in a way where the optimisations yield the best results.
Take \gls{cf} and \gls{dce}. \MPin{
	brief explanation?
}
Unfortunately, the order in which \gls{cf} and \gls{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
\MPin{figure here, or closer, not far away}
This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration\MP{of what}.\MP{cit?}
Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
\Cref{corr:swappable} justifies that any such order of compilation passes is valid with respect to security.
So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.

@@

\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
Besides sequential composition, there are two other compositions that are more at the language- than the compiler-level\MP{namely ...}.
Nevertheless, consider $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
As an example, $\src{Java}$ and $\obj{Kotlin}$ compile to $\trg{JVM BC}$.
Technically speaking, they both lower\MP{undefined verb. no need for synonyms here.} to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language, or rather, intermediate representation, performs a kind of {\em upper} composition.
Given a secure compiler from $\src{Java}$ to $\trg{JVM BC}$ and one from $\obj{Kotlin}$ to $\trg{JVM BC}$, it is easy to envision that one can engineer a combined compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$.
A practical example of this, which is, however, insecure, is {\tt Android Studio}~\cite{androidstudio}.
The proof of \Cref{thm:rtp} can be easily adapted to also works for this {\em upper} composition.
The corresponding theorem is not listed, since there is no general definition of this kind of composition.
\MPin{
	mmm need to think about this.
	this looks like discussion section material
}
\MPin{
	we could turn this into a : "if you expected upper adn lower, they are language composition ops, not compiler ones, so that's out of scope"
}

@@

\begin{figure}
  \begin{center}
    \begin{tikzpicture}
      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left] (progunopt) {\begin{lstlisting}[language=ml,basicstyle=\scriptsize\ttfamily]
let a = true in
if a then
  print "a"
else
  print "b"
\end{lstlisting}};
      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progunopt] (progoptcf) {\begin{lstlisting}[language=ml,basicstyle=\scriptsize\ttfamily]
if true then
  print "a"
else
  print "b"
\end{lstlisting}};
      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progoptcf] (progoptcfdce) {\begin{lstlisting}[language=ml,basicstyle=\scriptsize\ttfamily]
print "a"
\end{lstlisting}};% the following is a hack to get my syntax highlighting to work again \end{lstlisting}
      % edges
      \draw[->,very thick] (progunopt) edge[loop left,left] node {\gls{dce}} (progunopt);
      \draw[->,very thick] (progunopt) edge[sloped,above] node {\gls{cf}} (progoptcf);
      \draw[->,very thick] (progoptcf) edge[sloped,above] node {\gls{dce}} (progoptcfdce);
    \end{tikzpicture}
  \end{center}
  \caption{Example program where the level of optimisations differ for one pass of applying \gls{cf} and \gls{dce} in any order. %
  Every edge is a compilation pass and the label on the edge states what the pass does, i.e., \gls{cf} or \gls{dce}. %
  The source code in the nodes is a glorified compiler intermediate representation and the code gets more optimised towards the right hand side of the figure.}\label{fig:cfdceex}
\end{figure}

%\section{Case Study: Memory Safe and Cryptographic Constant Time Code}\label{sec:casestud}
\section{Case Study: Language Formalisations\pages{2}}\label{sec:casestud:defs}

This section defines programming languages that the secure compilers defined in the next section will make use of\MP{make use of -> use! 3 words when 1 is sufficient}.
To this end, this section defines the languages $\Ltms$, $\Ltrg$, $\Lms$, and $\Lscct$ which share many common elements (presented in \Cref{subsec:cs:defs}).
In this paper, $\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls{tms} (\Cref{??}).
However, not all $\Ltms$ programs are \gls{sms}.\MP{reworded. was imprecise}
That is, there are well--typed $\Ltms$ programs that perform an out--of--bounds access.
In contrast, $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls{ms} (\Cref{??}).
$\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability (\Cref{??}).
All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls{cct} to hold.

Even though it is possible to write programs that, e.g., return early in loops, which is trivially not \gls{cct}, they can be transformed to adhere to \gls{cct}~\cite{cauligi2019fact}.
\MPin{
	no, it's different. 
	there is no CCT because we asume it's out of the programmer's concers and since CT is arch-dependent, the compiler should preserve CT, not the programmer, via software
}

Language $\Lscct$ allows violating \gls{cct} by emitting events on, e.g., branching and division, that contain secrets (\Cref{??}).
However, $\Lscct$ also contains a countermeasure to enforce \gls{cct}, the language provides a way to read and write to a {\em model--specific register} that enables a ``\gls{cct}-mode''.
\MPin{ref to what real world stuff justifies this}

\subsection{Shared Language Definitions}\label{subsec:cs:defs}
\begin{gather*}
  \begin{aligned}
  \mi{(Expressions)}~&\expr\bnfdef\ x \mid \valueexpr \mid \lbinop{\expr[_{1}]}{\expr[_{2}]} \mid \lget{x}{\expr} \mid \lnew{\expr[_{1}]}{\expr[_{2}]} \mid \ldelete{x} \mid \lset{x}{\expr[_{1}]}{\expr[_{2}]} \\
    & \mid \llet{x}{\expr[_{1}]}{\expr[_{2}]}  \mid \lreturn{\expr} \mid \lcall{g}{\expr} \mid \lifz{\expr}{\expr[_{1}]}{\expr[_{2}]} \mid \labort
  \end{aligned}
  \\
  \begin{aligned}
  \mi{(Symbols)}~&\function\bnfdef\ \lfunction{g}{x}{e}
  &
  \mi{(Libraries)}~&\library\bnfdef\ \hole{\cdot} \mid \function,\library \\
  \mi{(Programs)}~&\prog{\library_{\ctx}}{\library_{\comp}}
  &
  \mi{(Substitutions)}~&\substvar\bnfdef \hole{\cdot} \mid \subst{\valueexpr}{x},\gamma
  \end{aligned}
\end{gather*}
\MPin{
	this latex had many issues.
	the \& were at different spots
	it used hspace to space things instead of \&
	and it did not use \& to align the following things on a line
	and it did not use multiple aligned into the gather
	(what's the point of gather if you don't have several aligned inside? use a single align* instead)
}

Above is a fraction of the shared syntax of all the programming languages of this paper.
\MP{why a fraction? what exprs are missing}
The syntax and semantics are mostly standard.\MP{and therefore omitted}
However, when allocating pointers\MP{arrays?} with $\lnew{\expr[_{1}]}{\expr[_{2}]}$, the expression $\expr[_{1}]$ determines the initial value of the objects\MP{cells?} and $\expr[_{2}]$ the overall size of an array.
The data associated to pointers resides in the dynamic state $\Omega$ that contains sandboxed heaps, one for the context and the other for the component.
Because of the sandboxing, locations pointing into heaps must be associated with a control--tag\MP{symbol} to know the heap they point into.
Other metadata in $\Omega$ entails a list of callable objects\MP{functions? objects are other things in PL} with information on wether they belong to the context or the component.
That way, when calling a callable object, the semantics can take care of emitting appropriate events that signal a context switch.
For the languages considered in this paper, no callable objects are first-class.
So, the only values are pairs, pointers, and natural numbers.
Another non-standard part of the languages is the concept of the poison-tag.
The poison-tag marks a location as freed ($\poisoned$) without removing the information associated to the freed pointer from $\Omega$.
This measure prevents the semantics from getting stuck when an already freed pointer is freed again.
Because of this, it is possible that programs like $\llet{x}{\lnew{0}{5}}{\ldelete{x};\ldelete{x}}$ reduce and emit a trace along the lines of $\ev{Alloc\ \loc\ \sandboxtag\ 5}\cdot\ev{Dealloc\ \loc\ \sandboxtag}\cdot\ev{Dealloc\ \loc\ \sandboxtag}$.

\begin{gather*}
  \begin{aligned}
  \mi{(Pre\text{-}Events)}~&\event[_{b}]\bnfdef \ev{Alloc}\ \loc\ \valueexpr \mid \ev{Dealloc}\ \loc \mid \ev{Get}\ \loc\ \valueexpr &\mid \ev{Set}\ \loc\ \valueexpr\ \valueexpr['] \mid \ev{Call}\ \comm\ g\ \valueexpr \mid \ev{Ret}\ \comm\ \valueexpr \mid \ev{Start} \mid \ev{End}\ \valueexpr \\
  \mi{(Events)}~&\event\bnfdef \emptyevent \mid \lightning \mid \tup{\event[_{b}]}{\sandboxtag} \\
  \end{aligned}
\end{gather*}
